### 2012-11-28 paper/paper.tex
Add skeleton for term paper

\documentclass[a4paper, twocolumns]{scrartcl}

\title{DNSKEY Management}
\author{Julien Perrochet \and Tobias Schlatter}
\date{December 12, 2012}

\begin{document}

\maketitle

\nocite{*}

\bibliographystyle{abbrv}
\bibliography{dnssec}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2012-12-01 paper/paper.tex
Add some introduction bullshit to paper

@@
-\documentclass[a4paper, twocolumns]{scrartcl}
+\documentclass[a4paper, twocolumn]{scrartcl}
 
@@
 
+\section{Introduction}
+%% TODO add some blah about management and penetration.
+
+In the following, every necessary operation for key management in
+DNSSEC is described as a reminder to the reader.
+
+\paragraph{Enabling DNSSEC} When enabling DNSSEC for a zone, the
+parent zone has to establish and sign a DS record with the newly
+created public-key.
+
+\paragraph{Key Rollover} RFC4641 \cite{RFC4641} suggests that the DNSKEY
+records for a zone should be changed somewhere between once every week
+up to once every month. While the detailed rollover procedure is
+irrelevant here, it is important to know, that the parent zone has to
+update the its DS record.
+
+\paragraph{Disabling DNSSEC} An operator might choose to discontinue
+DNSSEC for a zone. The DS records (or other means of signing DNSKEYs
+as we will see later) need to be removed.
+
 \nocite{*}
-
 \bibliographystyle{abbrv}

### 2012-12-02 paper/paper.tex
Proposed outline
Add command to verify char count
Write some stuff in paper (about DNSSEC...)

@@
-\documentclass[a4paper, twocolumn]{scrartcl}
+\documentclass[a4paper]{scrartcl}
+
+\usepackage{graphicx}
+
+%% Limit: 40000 chars
+%% check with `detex paper.tex | wc -c`
 
@@
 \date{December 12, 2012}
+
+\graphicspath{{../figures/}}
+
+\newcommand{\wbjp}{\protect\footnote{Written by Julien Perrochet}}
+\newcommand{\wbts}{\protect\footnote{Written by Tobias Schlatter}}
 
@@
 
+\section{DNSSEC Deployment\wbts}
+
+This section shall expose the current deployment status of DNSSEC.
+
+\subsection{Penetration}
+
+In figure~\ref{fig:dnssec-growth} you can see the development of
+deployed, production DNSSEC zones (for the heuristics applied to
+determine a production zone, see \cite{secspider, Osterweil09}).
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{dnssec-growth}
+  \caption{Number of deployed DNSSEC zones by
+    SecSpider \cite{secspider}}
+  \label{fig:dnssec-growth}
+\end{figure}
+
+\subsection{Trust Anchors}
+
+Contrary to the state at writing of \cite{Osterweil09}, the DNS
+root-zone is signed since July 15, 2010 \cite{root-dnssec} and hence
+serves as the global trust anchor.
+
+\section{Operations\wbjp}
+\subsection{3Rs}
+\subsection{Single admin}
+\subsection{Multi admin}
+\subsection{Disabling}
+
+\section{Case-Study: Switzerland (\texttt{ch.} Zone)\wbts}
+According to a query on \cite{secspider}, the \verb|ch.| TLD zone is signed and
+fully veryfiable from the root zone anchor. The validity period of a
+zone signing key is 37 days \cite{switch10}.
+
 \nocite{*}

### 2012-12-02 paper/paper.tex
Write more about DNSSEC penetration

@@
 \usepackage{graphicx}
+\usepackage{url}
 
@@
 deployed, production DNSSEC zones (for the heuristics applied to
-determine a production zone, see \cite{secspider, Osterweil09}).
+determine a production zone, see \cite{secspider, Osterweil09}). We
+can see that DNSSEC is gradually being deployed.
+
+Contrary to the state at writing of \cite{Osterweil09}, the DNS
+root-zone is signed since July 15, 2010 \cite{root-dnssec} and hence
+serves as the global trust anchor. However, as of this writing,
+data\footnote{\url{http://secspider.cs.ucla.edu/islands.html}}
+gathered by \cite{secspider} suggests that more than half of the
+deployed DNSSEC zones which are verifyable (i.e. are likely not to
+have a spoofed key, see \cite{Osterweil09} for details), do not yet
+have a fully linked certificate chain up to the root zone.
 
@@
 \subsection{Trust Anchors}
-
-Contrary to the state at writing of \cite{Osterweil09}, the DNS
-root-zone is signed since July 15, 2010 \cite{root-dnssec} and hence
-serves as the global trust anchor.
 

### 2012-12-03 paper/paper.tex
Some more about DNSSEC deployment and TARs

@@
-\documentclass[a4paper]{scrartcl}
+\documentclass[a4paper,twocolumn]{scrartcl}
 
 \usepackage{graphicx}
+%\usepackage{hyperref}
 \usepackage{url}
@@
 %% Limit: 40000 chars
-%% check with `detex paper.tex | wc -c`
+%% check with `detex paper.tex | wc -m`
 
@@
 
-This section shall expose the current deployment status of DNSSEC.
+This section will expose the current deployment status of DNSSEC and
+give the reader an idea about the current DNSSEC landscape before
+going into the specific DNSKEY managment issues. Please refer to
+section~\ref{sec:case-study} for more details on the situation of the
+\verb|ch.| zone.
 
-\subsection{Penetration}
+\subsection{Penetration in Production}
+In figure~\ref{fig:dnssec-growth} you can see the development of
+deployed, production DNSSEC zones. The data comes from SecSpider
+\cite{secspider} which crawls as many DNSSEC zones as possible,
+verifies whether it is properly signed and then applies a heuristics to
+determine whether it is actually a production zone. The goal is to
+rule out zones which are created for testing such as
+\url{bogussig.bogussig.test.jelte.nlnet.labs.nl}.
 
-In figure~\ref{fig:dnssec-growth} you can see the development of
-deployed, production DNSSEC zones (for the heuristics applied to
-determine a production zone, see \cite{secspider, Osterweil09}). We
-can see that DNSSEC is gradually being deployed.
+We can see that DNSSEC is gradually being deployed and the number of
+signed production zones grows rapidly (note the log-scale!). The rapid
+increase of crawled DNSSEC zones mid-2008, so interprets
+\cite{Osterweil09}, is due to a cache poisoning attack discovered in
+summer 2008 \cite{dnsVuln}. During our research, we were unfortunately
+unable to find an estimate of the number of (production) DNS zones in
+the internet to estimate the percentage of DNSSEC penetration.
 
@@
 deployed DNSSEC zones which are verifyable (i.e. are likely not to
-have a spoofed key, see \cite{Osterweil09} for details), do not yet
-have a fully linked certificate chain up to the root zone.
+have a spoofed key, see SecSpider in the next section or
+\cite{Osterweil09} for details), do not have a fully linked
+certificate chain up to the root zone.
 
-\begin{figure}
+\begin{figure*}
   \includegraphics[width=\linewidth]{dnssec-growth}
@@
   \label{fig:dnssec-growth}
-\end{figure}
+\end{figure*}
 
-\subsection{Trust Anchors}
+\subsection{Trust Anchor Repositories}
+Not always can the certificate chain be ensured from the root-zone
+down to a DNSSEC zone. Before the signing of the root-zone, it was
+obviously impossible to do such a thing, after the signing -- as
+formerly discussed data suggests -- there are still orphaned zones
+whose parent isn't properly signed.
+
+A Trust Anchor Repository (TAR) can be used in such a situation to
+supply resolvers with the trusted public key for a given zone during
+DNS resolution, therefore the name \emph{inline repository}. One of
+these repositories is the Interim Trust Anchor Repository (ITAR)
+formerly operated by IANA and discontinued since the signing of the
+root-zone \cite{itar}.
+
+The Internet Software Consortium (ISC) still runs an inline TAR
+\cite{iscDlv}, using the DNSSEC Lookaside Validation standard
+\cite{RFC5074}. When a resolver needs to check the key of a zone
+\verb|x| with this TAR, it makes a DNS request of type DLV to
+\url{x.dlv.isc.org}. If it succeeds, the DLV record contains the
+signature for the (hopefully matching) DNSKEY. The ISC inline TAR
+requires manual addition of keys to the repository.
+
+SecSpider \cite{secspider, Osterweil09} is another inline TAR, but
+uses a different approach for key validation: Rather then entering and
+signing public keys manually, it randomly queries DNSKEY entries from
+different point in the world simultaniously and only enters it into
+the TAR, if all the received keys match. This makes spoofing of a key
+very difficult due to the spatial distribution and the
+unpredictability of the queries.
+
+Another type of TAR resides at the resolver: A statically configured
+TAR stores keys locally at the resolver and hence needs not to issue an
+additional query for keys when a new DNS request arrives. The
+statically configured TAR may automatically poll keys asynchronously,
+or is maybe just a bunch of manually configured trusted keys. As an
+example of a statcially configured TAR, we have
+Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
+to a standard resolver, \cite{Osterweil09} which can poll keys from
+different locations such as web pages, but also also other
+vantage-enabled resolvers. Of course, the operater may also choose to
+add trusted keys manually.
 
@@
 \section{Case-Study: Switzerland (\texttt{ch.} Zone)\wbts}
+\label{sec:case-study}
 According to a query on \cite{secspider}, the \verb|ch.| TLD zone is signed and

### 2012-12-03 paper/paper.tex
Add more info about ch.

@@
 \label{sec:case-study}
-According to a query on \cite{secspider}, the \verb|ch.| TLD zone is signed and
-fully veryfiable from the root zone anchor. The validity period of a
-zone signing key is 37 days \cite{switch10}.
+
+\paragraph{Certificate Chain} According to a query on
+\cite{secspider}, the \verb|ch.| TLD zone is signed and fully
+veryfiable from the root zone anchor on downwards. The certificate
+chain can hence fully be established and no separate TARs are
+required.
+
+\paragraph{Rollover} The validity period of a zone signing key (for
+\verb|ch.|) is 37 days \cite{switch10}. For child zones,  DNSSEC and
+DNSKEY entries can be enabled and disabled through SWITCH's
+web-interface which is able to pull the keys automatically from the
+authoritative DNS servers. It is therefore relatively simple to enable
+DNSSEC for a \verb|x.ch.| zone, but using the web-interface might be
+cumbersome when having to deal with rollovers for a large number of
+domains.
 

### 2012-12-03 paper/paper.tex
Add more about situation in CH

@@
 veryfiable from the root zone anchor on downwards. The certificate
-chain can hence fully be established and no separate TARs are
-required.
+chain can hence fully be established and no separate TAR is required.
 
-\paragraph{Rollover} The validity period of a zone signing key (for
-\verb|ch.|) is 37 days \cite{switch10}. For child zones,  DNSSEC and
-DNSKEY entries can be enabled and disabled through SWITCH's
+\paragraph{Penetration} SWITCH states on their website that as of
+30. September 2012, 1'734'170 child-zones of \verb|ch.| have been
+registered\footnote{\url{https://www.nic.ch/reg/cm/wcm-page/statistics/index.html}}. Upon
+inquiry, SWITCH stated that there are around 370 secure delegations
+from the \verb|ch.| zone. This is less than $0.02\%$ penetration
+rate.
+
+\paragraph{Key Management} The validity period of a zone signing key
+(for \verb|ch.|) is 37 days \cite{switch10}. For child zones,  DNSSEC
+and DNSKEY entries can be enabled and disabled through SWITCH's 
 web-interface which is able to pull the keys automatically from the
-authoritative DNS servers. It is therefore relatively simple to enable
-DNSSEC for a \verb|x.ch.| zone, but using the web-interface might be
-cumbersome when having to deal with rollovers for a large number of
-domains.
+authoritative DNS servers. Partners (as Registrars are called in
+Switzerland) may also use EPP for key rollover.
+
+Therefore, it is relatively simple to enable DNSSEC for a \verb|x.ch.| 
+zone. While using the web-interface is certainly cumbersome when
+having to deal with rollovers for a large number of domains, any major
+provider may become a SWITCH Partner and can then switch over to EPP
+which solves this issue.
 

### 2012-12-03 paper/paper.tex
Minor improvements
Spellcheck

@@
 give the reader an idea about the current DNSSEC landscape before
-going into the specific DNSKEY managment issues. Please refer to
+going into the specific DNSKEY management issues. Please refer to
 section~\ref{sec:case-study} for more details on the situation of the
@@
 unable to find an estimate of the number of (production) DNS zones in
-the internet to estimate the percentage of DNSSEC penetration.
+the Internet to estimate the percentage of DNSSEC penetration.
 
@@
 gathered by \cite{secspider} suggests that more than half of the
-deployed DNSSEC zones which are verifyable (i.e. are likely not to
+deployed DNSSEC zones which are verifiable (i.e. are likely not to
 have a spoofed key, see SecSpider in the next section or
@@
 \cite{RFC5074}. When a resolver needs to check the key of a zone
-\verb|x| with this TAR, it makes a DNS request of type DLV to
-\url{x.dlv.isc.org}. If it succeeds, the DLV record contains the
+\verb|<X>| with this TAR, it makes a DNS request of type DLV to 
+\url{<X>.dlv.isc.org}. If it succeeds, the DLV record contains the
 signature for the (hopefully matching) DNSKEY. The ISC inline TAR
@@
 signing public keys manually, it randomly queries DNSKEY entries from
-different point in the world simultaniously and only enters it into
+different point in the world simultaneously and only enters it into
 the TAR, if all the received keys match. This makes spoofing of a key
@@
 or is maybe just a bunch of manually configured trusted keys. As an
-example of a statcially configured TAR, we have
+example of a statically configured TAR, we have
 Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
@@
 different locations such as web pages, but also also other
-vantage-enabled resolvers. Of course, the operater may also choose to
+vantage-enabled resolvers. Of course, the operator may also choose to
 add trusted keys manually.
@@
 \subsection{3Rs}
+%% Talk about EPP or push burden to Case-Study
 \subsection{Single admin}
@@
 \cite{secspider}, the \verb|ch.| TLD zone is signed and fully
-veryfiable from the root zone anchor on downwards. The certificate
+verifiable from the root zone anchor on downwards. The certificate
 chain can hence fully be established and no separate TAR is required.
@@
 
-Therefore, it is relatively simple to enable DNSSEC for a \verb|x.ch.| 
+Therefore, it is relatively simple to enable DNSSEC for a \verb|<X>.ch.| 
 zone. While using the web-interface is certainly cumbersome when

### 2012-12-03 paper/paper.tex
Add paragraph about cost of DNSSEC @ SWITCH
Add abstract

@@
 \maketitle
+
+\begin{abstract}
+  In this paper we recapitutale the current deployment status of
+  DNSSEC and go into the main issues when managing its keys. It
+  outlines the paper ``Interadministrative Challenges in Managing
+  DNSKEYs'' and updates some of its contents especially with respect
+  to the signing of the root zone on July 15, 2012. We will see that
+  DNSSEC penetration in the Internet's zones is still low but that the
+  necessary foundations have been lain for full deployment.
+\end{abstract}
 
@@
 
+\paragraph{Cost} DNSSEC secure delegation is currently provided by
+SWITCH with no additional cost to normal DNS delegation.
+
 \nocite{*}

### 2012-12-03 paper/paper.tex
Nicely typeset abstract

--- 2012-12-03 paper/paper.tex
@@
 \usepackage{url}
+\usepackage{abstract}
 
@@
 
+\twocolumn[
 \maketitle
+\begin{onecolabstract}
+  In this paper we recapitutale the current deployment status of
+  DNSSEC and go into the main issues when managing its keys. It
+  outlines the paper ``Interadministrative Challenges in Managing
+  DNSKEYs'' \cite{Osterweil09} and updates some of its contents
+  especially with respect to the signing of the root zone on July
+  15, 2012. We will see that DNSSEC penetration in the Internet's
+  zones is still low but that the necessary foundations have been lain
+  for full deployment.
+\end{onecolabstract}
+\vspace{\baselineskip}
+]
 

### 2012-12-04 paper/paper.tex
singl admin zone and first blabla

--- 2012-12-03 paper/paper.tex
@@
 give the reader an idea about the current DNSSEC landscape before
-going into the specific DNSKEY management issues. Please refer to
+going into the specific DNSKEY managment issues. Please refer to
 section~\ref{sec:case-study} for more details on the situation of the
@@
 unable to find an estimate of the number of (production) DNS zones in
-the Internet to estimate the percentage of DNSSEC penetration.
+the internet to estimate the percentage of DNSSEC penetration.
 
@@
 gathered by \cite{secspider} suggests that more than half of the
-deployed DNSSEC zones which are verifiable (i.e. are likely not to
+deployed DNSSEC zones which are verifyable (i.e. are likely not to
 have a spoofed key, see SecSpider in the next section or
@@
 \cite{RFC5074}. When a resolver needs to check the key of a zone
-\verb|<X>| with this TAR, it makes a DNS request of type DLV to 
-\url{<X>.dlv.isc.org}. If it succeeds, the DLV record contains the
+\verb|x| with this TAR, it makes a DNS request of type DLV to
+\url{x.dlv.isc.org}. If it succeeds, the DLV record contains the
 signature for the (hopefully matching) DNSKEY. The ISC inline TAR
@@
 signing public keys manually, it randomly queries DNSKEY entries from
-different point in the world simultaneously and only enters it into
+different point in the world simultaniously and only enters it into
 the TAR, if all the received keys match. This makes spoofing of a key
@@
 or is maybe just a bunch of manually configured trusted keys. As an
-example of a statically configured TAR, we have
+example of a statcially configured TAR, we have
 Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
@@
 different locations such as web pages, but also also other
-vantage-enabled resolvers. Of course, the operator may also choose to
+vantage-enabled resolvers. Of course, the operater may also choose to
 add trusted keys manually.
@@
 \section{Operations\wbjp}
+This section covers the issues and quirks related to the management of DNSSEC.
 \subsection{3Rs}
-%% Talk about EPP or push burden to Case-Study
-\subsection{Single admin}
-\subsection{Multi admin}
-\subsection{Disabling}
+Regarding the management of the internet namespace, three entities can generally be considered:
+
+\begin{itemize}
+\item the registry;
+\item the registrant;
+\item the registrar.
+\end{itemize}
+
+The registries are the entities serving the records for a zone.
+For example, Switch is the registry for the \verb|ch.| zone.
+Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch| is the school in itself, for example.
+
+These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a domain name and an IP address.
+
+In Switzerland, for example, Switch plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from Switch.
+Registries will generally delegate part or whole of the namespace registration managing to registrars.
+
+Registrants are also free to play the registry and registrant roles for their own sub-domains: groups, laboratories and associations within EPFL will obtain an \verb|epfl.ch| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch| subdomain, for example.
+
+The entities behind the three R's are the exact same wether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data actually changed.
+
+\subsection{Signing a zone}
+Whenever an entity wishes to sign its zone, it should ideally turn to its already signed parent zone, in order to create a continuous thrust chain.
+Said parent should then serve a DS record containing the child zone's key signature.
+
+Signing new zones will have different administrative implications depending on whether the parent zone is in the same administrative entity or not: e.g., \verb|www.epfl.ch| is managed by the same entity as \verb|epfl.ch|, so signing the \verb|www| sub-domain will imply a different (and very likely easier) procedure than signing \verb|epfl.ae|, EPFL's domain for its middle-east campus.
+
+\subsubsection{Single administrative domain}
+Basically, once an organisation has its main domain signed, it is relatively easy for any sub-entity within said organisation to get a sub-domain signed, and to coordinate any operations relative to DNSSEC like key roll-over: staying with our EPFL example, it can suffice to contact the IT-department and discuss about the procedure with them\footnote{EPFL does not actually use DNSSEC, but we use the school as an example for the sake of simplicity.}.
+The bottom line is that it is relatively easy to handle the additional work and information updates required by DNSSEC as long as every concerned party lies within the same organization.
+\subsubsection{Multiple administrative domain}
+\subsection{Disabling DNSSEC}
 
@@
 \cite{secspider}, the \verb|ch.| TLD zone is signed and fully
-verifiable from the root zone anchor on downwards. The certificate
-chain can hence fully be established and no separate TAR is required.
+veryfiable from the root zone anchor on downwards. The certificate
+chain can hence fully be established and no separate TARs are
+required.
 
-\paragraph{Penetration} SWITCH states on their website that as of
-30. September 2012, 1'734'170 child-zones of \verb|ch.| have been
-registered\footnote{\url{https://www.nic.ch/reg/cm/wcm-page/statistics/index.html}}. Upon
-inquiry, SWITCH stated that there are around 370 secure delegations
-from the \verb|ch.| zone. This is less than $0.02\%$ penetration
-rate.
-
-\paragraph{Key Management} The validity period of a zone signing key
-(for \verb|ch.|) is 37 days \cite{switch10}. For child zones,  DNSSEC
-and DNSKEY entries can be enabled and disabled through SWITCH's 
+\paragraph{Rollover} The validity period of a zone signing key (for
+\verb|ch.|) is 37 days \cite{switch10}. For child zones,  DNSSEC and
+DNSKEY entries can be enabled and disabled through SWITCH's
 web-interface which is able to pull the keys automatically from the
-authoritative DNS servers. Partners (as Registrars are called in
-Switzerland) may also use EPP for key rollover.
-
-Therefore, it is relatively simple to enable DNSSEC for a \verb|<X>.ch.| 
-zone. While using the web-interface is certainly cumbersome when
-having to deal with rollovers for a large number of domains, any major
-provider may become a SWITCH Partner and can then switch over to EPP
-which solves this issue.
+authoritative DNS servers. It is therefore relatively simple to enable
+DNSSEC for a \verb|x.ch.| zone, but using the web-interface might be
+cumbersome when having to deal with rollovers for a large number of
+domains.
 

### 2012-12-04 paper/paper.tex
typo correction and more stuff.

--- 2012-12-03 paper/paper.tex
@@
 \section{Operations\wbjp}
+This section covers the issues and quirks related to the management of DNSSEC.
 \subsection{3Rs}
 %% Talk about EPP or push burden to Case-Study
-\subsection{Single admin}
-\subsection{Multi admin}
-\subsection{Disabling}
+Regarding the management of the internet namespace, three entities can generally be considered:
+
+\begin{itemize}
+\item the registry;
+\item the registrant;
+\item the registrar.
+\end{itemize}
+
+The registries are the entities serving the records for a zone.
+For example, Switch is the registry for the \verb|ch.| zone.
+Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch| is the school in itself, for example.
+
+These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a domain name and an IP address.
+
+In Switzerland, for example, Switch plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from Switch.
+Registries will generally delegate part or whole of the namespace registration managing to registrars.
+
+Registrants are also free to play the registry and registrant roles for their own sub-domains: groups, laboratories and associations within EPFL will obtain an \verb|epfl.ch| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch| subdomain, for example.
+
+The entities behind the three R's are the exact same wether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data actually changed.
+
+\subsection{Signing a zone}
+Whenever an entity wishes to sign its zone, it should ideally turn to its already signed parent zone, in order to create a continuous trust chain, as said parent will then serve a DS record containing the child zone's key signature.
+
+Signing new zones will have different administrative implications depending on whether the parent zone is in the same administrative entity or not: e.g., \verb|www.epfl.ch| is managed by the same entity as \verb|epfl.ch|, so signing the \verb|www| sub-domain will imply a different (and very likely easier) procedure than signing \verb|epfl.ae|, EPFL's domain for its middle-east campus.
+
+\subsubsection{Single administrative domain}
+Basically, once an organisation has its main domain signed, it is relatively easy for any sub-entity within said organisation to get a sub-domain signed, and to coordinate any operations relative to DNSSEC like key roll-over: staying with our EPFL example, it can suffice to contact the IT-department and discuss about the procedure with them\footnote{EPFL does not actually use DNSSEC, but we use the school as an example for the sake of simplicity.}.
+The bottom line is that it is relatively easy to handle the additional work and information updates required by DNSSEC as long as every concerned party lies within the same organization.
+\subsubsection{Multiple administrative domain}
+DNSSEC's true challenges arise when we begin to consider several administrative entities (the three R's, mainly).
+As stated earlier, once the \verb|epfl.ch| zone is signed it is easy for the school to sign \verb|www.epfl.ch|.
+However, if the domain to be signed was \verb|epfl.ch| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
+
+While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant for such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are.
+\subsection{Disabling DNSSEC}
 

### 2012-12-04 paper/paper.tex
Introduction to ch. zone

@@
 
-\section{Case-Study: Switzerland (\texttt{ch.} Zone)\wbts}
+\section{Case-Study: \texttt{ch.} Zone\wbts}
 \label{sec:case-study}
 
-\paragraph{Certificate Chain} According to a query on
+The \verb|ch.| zone is managed by SWITCH, the institution which also
+employs the Swiss educational network. In this section we briefly
+point out the current deployment status of DNSSEC and how SWITCH
+handles the discussed key management issues.
+
+\paragraph{Certificate Chain} According to a query on SecSpider
 \cite{secspider}, the \verb|ch.| TLD zone is signed and fully

### 2012-12-04 paper/paper.tex
Improve introduction

@@
 \section{Introduction}
-%% TODO add some blah about management and penetration.
+DNSSEC is gradually being deployed in the Internet. However, managing
+the cryptographic keys required for the secure delegation provided by
+DNSSEC is not always entirely trivial.
 
+In this paper we will first remind of the prinicpal necessary
+operations for key management in DNSSEC. Then we will give the reader
+an overview of the current status of DNSSEC deployment in the Internet
+to put the in-detail discussion about key management into the right
+context.
+
+Finally, we will shortly outline the DNSSEC situation in Switzerland
+(i.e. the \verb|ch.|  and how its authorities handle DNSKEY management
+issues.
+
+\subsection{Operations}
 In the following, every necessary operation for key management in
-DNSSEC is described as a reminder to the reader.
+DNSSEC is described as a reminder to the reader. This is by no means
+intended to give an introduction to DNSSEC but rather to remind the
+reader about specific details particularily relevant to this paper.
 

### 2012-12-04 paper/paper.tex
Spellcheck

@@
 \begin{onecolabstract}
-  In this paper we recapitutale the current deployment status of
+  In this paper we recapitulate the current deployment status of
   DNSSEC and go into the main issues when managing its keys. It
@@
 
-In this paper we will first remind of the prinicpal necessary
+In this paper we will first remind of the principal necessary
 operations for key management in DNSSEC. Then we will give the reader
@@
 intended to give an introduction to DNSSEC but rather to remind the
-reader about specific details particularily relevant to this paper.
+reader about specific details particularly relevant to this paper.
 
@@
 irrelevant here, it is important to know, that the parent zone has to
-update the its DS record.
+update its DS record.
 
@@
 %% Talk about EPP or push burden to Case-Study
-Regarding the management of the internet namespace, three entities can generally be considered:
+Regarding the management of the Internet namespace, three entities can generally be considered:
 
@@
 
-The entities behind the three R's are the exact same wether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data actually changed.
+The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data actually changed.
 
@@
 \subsubsection{Single administrative domain}
-Basically, once an organisation has its main domain signed, it is relatively easy for any sub-entity within said organisation to get a sub-domain signed, and to coordinate any operations relative to DNSSEC like key roll-over: staying with our EPFL example, it can suffice to contact the IT-department and discuss about the procedure with them\footnote{EPFL does not actually use DNSSEC, but we use the school as an example for the sake of simplicity.}.
+Basically, once an organization has its main domain signed, it is relatively easy for any sub-entity within said organization to get a sub-domain signed, and to coordinate any operations relative to DNSSEC like key rollover: staying with our EPFL example, it can suffice to contact the IT-department and discuss about the procedure with them\footnote{EPFL does not actually use DNSSEC, but we use the school as an example for the sake of simplicity.}.
 The bottom line is that it is relatively easy to handle the additional work and information updates required by DNSSEC as long as every concerned party lies within the same organization.

### 2012-12-04 paper/paper.tex
Add chain of trust to introduction, yet to fill
latest.

--- 2012-12-04 paper/paper.tex
@@
 \subsection{Signing a zone}
-Whenever an entity wishes to sign its zone, it should ideally turn to its already signed parent zone, in order to create a continuous trust chain, as said parent will then serve a DS record containing the child zone's key signature.
+Whenever an entity wishes to sign its zone, it should ideally turn to its already signed parent zone, in order to create a continuous thrust chain.
+Said parent should then serve a DS record containing the child zone's key signature.
 
@@
 \subsubsection{Multiple administrative domain}
-DNSSEC's true challenges arise when we begin to consider several administrative entities (the three R's, mainly).
-As stated earlier, once the \verb|epfl.ch| zone is signed it is easy for the school to sign \verb|www.epfl.ch|.
-However, if the domain to be signed was \verb|epfl.ch| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
-
-While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant for such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are.
 \subsection{Disabling DNSSEC}
 
-\section{Case-Study: \texttt{ch.} Zone\wbts}
+\section{Case-Study: Switzerland (\texttt{ch.} Zone)\wbts}
 \label{sec:case-study}
 
-The \verb|ch.| zone is managed by SWITCH, the institution which also
-employs the Swiss educational network. In this section we briefly
-point out the current deployment status of DNSSEC and how SWITCH
-handles the discussed key management issues.
-
-\paragraph{Certificate Chain} According to a query on SecSpider
+\paragraph{Certificate Chain} According to a query on
 \cite{secspider}, the \verb|ch.| TLD zone is signed and fully

### 2012-12-04 paper/paper.tex

--- 2012-12-04 paper/paper.tex
@@
 
-In this paper we will first remind of the principal necessary
-operations for key management in DNSSEC. Then we will give the reader
-an overview of the current status of DNSSEC deployment in the Internet
-to put the in-detail discussion about key management into the right
+In this paper we will first remind of the concepts and operations in
+DNSSEC relevant for key management. Then we will give the reader an
+overview of the current status of DNSSEC deployment in the Internet to
+put the in-detail discussion about key management into the right
 context.
@@
 issues.
+
+\subsection{Chain of Trust}
+%% TODO add
 
@@
 \subsubsection{Multiple administrative domain}
-DNSSEC's true challenges arise when we begin to consider several administrative entities (the three R's, mainly).
-As stated earlier, once the \verb|epfl.ch| zone is signed it is easy for the school to sign \verb|www.epfl.ch|.
-However, if the domain to be signed was \verb|epfl.ch| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
-
-While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant for such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are.
 \subsection{Disabling DNSSEC}

### 2012-12-04 paper/paper.tex
oups, deleted some stuff, putting it back.

--- 2012-12-04 paper/paper.tex
@@
 \begin{itemize}
-\item the registry;
-\item the registrant;
-\item the registrar.
+\item a registry;
+\item a registrant;
+\item a registrar.
 \end{itemize}
@@
 For example, Switch is the registry for the \verb|ch.| zone.
-Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch| is the school in itself, for example.
+Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch| is the school in itself (its IT-department to be precise \footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
 
@@
 In Switzerland, for example, Switch plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from Switch.
-Registries will generally delegate part or whole of the namespace registration managing to registrars.
+Registries will generally delegate part or whole of the namespace registration managing to registrars\footnote{For example, one can also buy a .ch domain from \url{http://gandi.net}}
 
-Registrants are also free to play the registry and registrant roles for their own sub-domains: groups, laboratories and associations within EPFL will obtain an \verb|epfl.ch| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch| subdomain, for example.
+Registrants are also free to play the registry and registrar roles for their own sub-domains: groups, laboratories and associations within EPFL may apply for an \verb|epfl.ch| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch| subdomain, for example.
 
-The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data actually changed.
+The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data was actually updated.
 
 \subsection{Signing a zone}
-Whenever an entity wishes to sign its zone, it should ideally turn to its already signed parent zone, in order to create a continuous trust chain, as said parent will then serve a DS record containing the child zone's key signature.
+Whenever an entity wishes to sign its zone, it should ideally turn to its already signed parent zone, in order to create a continuous trust chain.
+Said parent should then serve a DS record containing the child zone's key signature.
 
@@
 While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant for such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are.
+
 \subsection{Disabling DNSSEC}

### 2012-12-04 paper/paper.tex
outsourcing stuff

@@
 
-While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant for such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are.
+While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as Switch's servers do not reflect the update.
+
+As suggested by some RFC\footnote{RFC 4641}, key updates should  happen at least every few months.
+Hence, it is crucial that a streamlined and reactive process is available for child-zones to notify their parent-zone when required.
+In the three R's setup, this means that the registry (or the registrars) must provide means for registrants to easily inform them of any changes regarding their DNSSEC keys.
+%TODO process figure with Registrant -> Registrar -> Registry ?
+
+Furthermore, as these communications concern the DNSSEC setup, security must be guaranteed: if a registry's update tools allow an attacker to update the key signatures, the whole point of DNSSEC is moot.
+
+\paragraph*{Entity Roles Overlap} Another issue arises when we consider the fact that registrants may outsource part or whole of the hosting burden, like:
+
+\begin{description}
+\item[Key Signing] Services that will handle the zone signing and key rollovers are available\footnote{like this registrar: \url{http://registrars.nominet.org.uk/registration-and-domain-management/dns/dnssec-signing-service}}.
+\item[Hosting] Countless hosting services are available, many of which are proposed by actual registrars.
+Some of them can also propose DNSSEC to their customers, in addition to the standard DNS service they also provide.
+\end{description}
+
+While having the registrars also handle DNSSEC will ease managing key rollovers and updates, it creates another problem: registrants, who generally are the registrars or registry's customers, might very well decide to switch from one service provider to another.
+
+If standard DNS simply provides zone transfers as a mean enabling such migrations, DNSSEC must be treated differently.
+Whether a new private-public key pair is generated or whether the existing pair is transferred, new hurdles are introduced:
+
+\begin{itemize}
+\item[-] The transferred private key is not totally \emph{private} anymore, as now both the old and new service provider know it;
+\item[-] Newly generated keys will require an update of the parent zone, and this might happen outside of normally planned rollover periods.
+\end{itemize}
+
+Another issue is that hosting and key signing services that are not registrars don't have to comply to any of ICANN's rules, increasing the difficulty to enforce common procedures, as none of these services is contractually bound to do so.
 

### 2012-12-04 paper/paper.tex
tea time.

@@
 
+\subsection{Computing power}
+DNSSEC does not only require a higher coordination among the three R's, it also increases the strain on the computing infrastructure of a registrar when the size and quantity of the  zones it manages grow: updates to a zone or key rollovers require cryptographic signing operations that are orders of magnitude more expensive than simply updating the DNS record.
+Even with improvements like incremental signing, the computational requirements are significant and must be considered when a registrar whishes to enable DNSSEC.
+%TODO find good ref on this.
+
 \subsection{Disabling DNSSEC}

### 2012-12-05 paper/paper.tex
break

@@
 \subsection{Disabling DNSSEC}
+The final and somewhat unintuitive managerial issue with DNSSEC is what needs to be done once a registrant wishes to turn it off.
 
+In such a case, it is absolutely insufficient to only turn off the service on the devices serving the aforementioned  zone: in doing so, resolvers won't be able to distinguish between spoofed DNS replies that contain non-DNSSEC data and legitimate replies that simply stopped using DNSSEC.
+
+For
 \section{Case-Study: \texttt{ch.} Zone\wbts}

### 2012-12-05 paper/paper.tex
dnssec off done

--- 2012-12-04 paper/paper.tex
@@
 \subsection{Disabling DNSSEC}
+The final and somewhat unintuitive managerial issue with DNSSEC is what needs to be done once a zone administrator wishes to turn it off.
+
+In such a case, it is absolutely insufficient to only disable the service on the devices serving the aforementioned  zone: in doing so, resolvers won't be able to distinguish between spoofed DNS replies that contain non-DNSSEC data and legitimate replies that simply stopped using DNSSEC.
+
+For DNSSEC's shutdown to be clean and complete, the parent zone and any TAR must be cleaned from whatever DS record they contain for the zone which is being withdrawn from DNSSEC.
+
+Again, this teardown must happen in a synchronous way, as resolvers will perceive the domain to be \emph{broken} as long as all involved parties are in a incoherent state.
 

### 2012-12-05 paper/paper.tex
trust chains.

@@
 
-\subsection{Chain of Trust}
-%% TODO add
+\subsection{Chain of Trust\wbjp}
+The goal of DNSSEC is to allow resolvers -- any person or process querying a DNS server -- to ensure the authenticity of the reply.
+Namely, in standard DNS, nothing guarantees the resolvers that his query's answer has not been modified or really originated from a legitimate server.
+
+DNSSEC permits zone operators to \emph{sign} the entries they serve to a resolver, enabling the latter to verify that what he receives is authentic.
+
+This is achieved by establishing a trust chain between something known and the domain that is being queried: e.g., I trust this person to tell me the truth and to be who it claims to be, because a friend I trust is trusting this person.
+
+DNSSEC trust chains are similar to the ones used in SSL, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public-private key pair with additional metadata.}
+
+Hence, in the real world, the \emph{friend} we trust is the so-called root authority -- the root domain in DNSSEC -- of whom we know the public key in advance.
+
+By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in \ref{fig:trust-chain}.
+This way, a resolver querying EPFL's DNS server will be able to verify the received query answer against the known root-authority in the following way:
+
+\begin{figure*}
+\center
+  \includegraphics[width=\linewidth]{trust-chain}
+  \caption{Example of a trust-hierarchy. Here, the resolver can establish a trust chain between the Root and the ucla.edu zone.}
+  \label{fig:trust-chain}
+\end{figure*}
 
@@
 
-Again, this teardown must happen in a synchronous way, as resolvers will perceive the domain to be \emph{broken} as long as all involved parties are in a incoherent state.
+Again, this teardown must happen in a synchronous way -- requiring additional effort with regard to DNS -- as resolvers will perceive the domain to be \emph{broken} as long as all involved parties are in a incoherent state.
 

### 2012-12-05 paper/paper.tex
trust chain progress. may need to finish it more nicely

@@
 
-By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in \ref{fig:trust-chain}.
-This way, a resolver querying EPFL's DNS server will be able to verify the received query answer against the known root-authority in the following way:
+By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in fig.~\ref{fig:trust-chain}.
+This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch| will be able to verify the received query answer against the known root-authority in the following way:
+
+\begin{enumerate}
+\item Verify the reply's authenticity using \verb|lacal.epfl.ch|'s public key, obtained from \verb|epfl.ch|'s DNS server ;
+\item Verify the query -- containing \verb|lacal.epfl.ch|'s public key -- returned by \verb|epfl.ch| with its public key, retrieved from the \verb|.ch| DNS server;
+\item Again, verify \verb|.ch|'s responses by getting its public key from the root domain.
+\item Finally, the root domain's responses can be verified with its already known public key.
+\end{enumerate}
 

### 2012-12-05 paper/paper.tex
finshed trust chain descr.

@@
 Finally, we will shortly outline the DNSSEC situation in Switzerland
-(i.e. the \verb|ch.|  and how its authorities handle DNSKEY management
+(i.e. the \verb|ch.| zone and how its authorities handle DNSKEY management
 issues.
@@
 
-DNSSEC trust chains are similar to the ones used in SSL, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public-private key pair with additional metadata.}
-
-Hence, in the real world, the \emph{friend} we trust is the so-called root authority -- the root domain in DNSSEC -- of whom we know the public key in advance.
+In the DNS world, the \emph{friend} we trust is the so-called root authority -- the root domain in DNSSEC -- of whom we know the public key in advance.
 
@@
 \end{enumerate}
+
+DNSSEC's trust hierarchy is similar to the SSL Certificate infrastructure, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public key with additional metadata.}.
+However, DNSSEC theoretically only has one single top-authority\footnote{The use of TAR's should remain temporary, as in an ideal setup DNSSEC could work without them.}, while the SSL Certificate hierarchy disposes of many.
 

### 2012-12-05 paper/paper.tex
minor corrections.

@@
 
-\subsection{Operations}
+\subsection{Operations\wbts}
 In the following, every necessary operation for key management in
@@
 %% Talk about EPP or push burden to Case-Study
-Regarding the management of the Internet namespace, three entities can generally be considered:
+Regarding the management of the Internet namespace, three entities can generally be considered, namely:
 
 \begin{itemize}
-\item a registry;
-\item a registrant;
-\item a registrar.
+\item \textbf{R}egistries;
+\item \textbf{R}egistrants;
+\item \textbf{R}egistrars.
 \end{itemize}

### 2012-12-05 paper/paper.tex
initials instead of footnotes for paragraph authors.

@@
 \title{DNSKEY Management}
-\author{Julien Perrochet \and Tobias Schlatter}
+\author{Julien Perrochet [JP] \and Tobias Schlatter [TS]}
 \date{December 12, 2012}
@@
 
-\newcommand{\wbjp}{\protect\footnote{Written by Julien Perrochet}}
-\newcommand{\wbts}{\protect\footnote{Written by Tobias Schlatter}}
+\newcommand{\wbjp}{\textsuperscript{JP}}
+\newcommand{\wbts}{\textsuperscript{TS}}
 
@@
 
-In the DNS world, the \emph{friend} we trust is the so-called root authority -- the root domain in DNSSEC -- of whom we know the public key in advance.
+In the DNS world, the \emph{friend} we trust is the so-called root authority -- the root domain -- of whom we know the public key in advance.
 
@@
 up to once every month. While the detailed rollover procedure is
-irrelevant here, it is important to know, that the parent zone has to
-update its DS record.
+irrelevant here, it is important to know that the parent zone has to
+update its DS record regularly.
 
@@
 \cite{secspider} which crawls as many DNSSEC zones as possible,
-verifies whether it is properly signed and then applies a heuristics to
+verifies whether they are properly signed and then applies a heuristic to
 determine whether it is actually a production zone. The goal is to
@@
 The Internet Software Consortium (ISC) still runs an inline TAR
-\cite{iscDlv}, using the DNSSEC Lookaside Validation standard
-\cite{RFC5074}. When a resolver needs to check the key of a zone
+\cite{iscDlv}, using the DNSSEC Lookaside Validation standard~\cite{RFC5074}. When a resolver needs to check the key of a zone
 \verb|<X>| with this TAR, it makes a DNS request of type DLV to 
@@
 
+Enabling DNSSEC for a zone not being a \emph{do once and forget} operation, but instead requires additional coordination among the three R's, let us see what the managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.
+
 \subsubsection{Single administrative domain}
@@
 The bottom line is that it is relatively easy to handle the additional work and information updates required by DNSSEC as long as every concerned party lies within the same organization.
+
 \subsubsection{Multiple administrative domain}
@@
 
-As suggested by some RFC\footnote{RFC 4641}, key updates should  happen at least every few months.
+As suggested by some RFC\cite{RFC4641} key updates should  happen at least every few months.
 Hence, it is crucial that a streamlined and reactive process is available for child-zones to notify their parent-zone when required.

### 2012-12-05 paper/paper.tex
Spellcheck paper

@@
 
-By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in fig.~\ref{fig:trust-chain}.
+By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in figure~\ref{fig:trust-chain}.
 This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch| will be able to verify the received query answer against the known root-authority in the following way:
@@
 DNSSEC does not only require a higher coordination among the three R's, it also increases the strain on the computing infrastructure of a registrar when the size and quantity of the  zones it manages grow: updates to a zone or key rollovers require cryptographic signing operations that are orders of magnitude more expensive than simply updating the DNS record.
-Even with improvements like incremental signing, the computational requirements are significant and must be considered when a registrar whishes to enable DNSSEC.
+Even with improvements like incremental signing, the computational requirements are significant and must be considered when a registrar wishes to enable DNSSEC.
 %TODO find good ref on this.

### 2012-12-05 paper/paper.tex
coherent dotting for domains/zones.

--- 2012-12-05 paper/paper.tex
@@
 By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in fig.~\ref{fig:trust-chain}.
-This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch| will be able to verify the received query answer against the known root-authority in the following way:
+This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch.| will be able to verify the received query answer against the known root-authority in the following way:
 
 \begin{enumerate}
-\item Verify the reply's authenticity using \verb|lacal.epfl.ch|'s public key, obtained from \verb|epfl.ch|'s DNS server ;
-\item Verify the query -- containing \verb|lacal.epfl.ch|'s public key -- returned by \verb|epfl.ch| with its public key, retrieved from the \verb|.ch| DNS server;
-\item Again, verify \verb|.ch|'s responses by getting its public key from the root domain.
+\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s public key, obtained from \verb|epfl.ch.|'s DNS server ;
+\item Verify the query -- containing \verb|lacal.epfl.ch.|'s public key -- returned by \verb|epfl.ch.| with its public key, retrieved from the \verb|ch.| DNS server;
+\item Again, verify \verb|ch.|'s responses by getting its public key from the root domain.
 \item Finally, the root domain's responses can be verified with its already known public key.
@@
 Regarding the management of the Internet namespace, three entities can generally be considered, namely:
-
 \begin{itemize}
@@
 For example, Switch is the registry for the \verb|ch.| zone.
-Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch| is the school in itself (its IT-department to be precise \footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
+Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department to be precise \footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
 
@@
 
-Registrants are also free to play the registry and registrar roles for their own sub-domains: groups, laboratories and associations within EPFL may apply for an \verb|epfl.ch| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch| subdomain, for example.
+Registrants are also free to play the registry and registrar roles for their own sub-domains: groups, laboratories and associations within EPFL may apply for an \verb|epfl.ch.| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch.| subdomain, for example.
 
@@
 
-Signing new zones will have different administrative implications depending on whether the parent zone is in the same administrative entity or not: e.g., \verb|www.epfl.ch| is managed by the same entity as \verb|epfl.ch|, so signing the \verb|www| sub-domain will imply a different (and very likely easier) procedure than signing \verb|epfl.ae|, EPFL's domain for its middle-east campus.
+Signing new zones will have different administrative implications depending on whether the parent zone is in the same administrative entity or not: e.g., \verb|www.epfl.ch.| is managed by the same entity as \verb|epfl.ch.|, so signing the \verb|www| sub-domain will imply a different (and very likely easier) procedure than signing \verb|epfl.ae|, EPFL's domain for its middle-east campus.
 
@@
 DNSSEC's true challenges arise when we begin to consider several administrative entities (the three R's, mainly).
-As stated earlier, once the \verb|epfl.ch| zone is signed it is easy for the school to sign \verb|www.epfl.ch|.
-However, if the domain to be signed was \verb|epfl.ch| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
+As stated earlier, once the \verb|epfl.ch.| zone is signed it is easy for the school to sign \verb|www.epfl.ch.|.
+However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
 
-While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as Switch's servers do not reflect the update.
+While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as Switch's servers do not reflect the update.
 

### 2012-12-06 paper/paper.tex
new figure.
Fix wrong reference name
Add some comments, some grammatical corrections, spellcheck

@@
   In this paper we recapitulate the current deployment status of
-  DNSSEC and go into the main issues when managing its keys. It
-  outlines the paper ``Interadministrative Challenges in Managing
-  DNSKEYs'' \cite{Osterweil09} and updates some of its contents
-  especially with respect to the signing of the root zone on July
-  15, 2012. We will see that DNSSEC penetration in the Internet's
-  zones is still low but that the necessary foundations have been lain
+  DNSSEC and go into the main issues when managing its keys. This
+  paper outlines the paper ``Interadministrative Challenges in
+  Managing DNSKEYs'' \cite{Osterweil09} and updates some of its
+  contents, especially with respect to the signing of the root zone on
+  July 15, 2012. We will see that DNSSEC penetration in the Internet's 
+  zones is still low, but that the necessary foundations have been lain
   for full deployment.
@@
 In this paper we will first remind of the concepts and operations in
-DNSSEC relevant for key management. Then we will give the reader an
-overview of the current status of DNSSEC deployment in the Internet to
-put the in-detail discussion about key management into the right
+DNSSEC which are relevant for key management. Then we will give the
+reader an overview of the current status of DNSSEC deployment in the
+Internet to put the discussion about key management into the right
 context.
@@
 Finally, we will shortly outline the DNSSEC situation in Switzerland
-(i.e. the \verb|ch.| zone and how its authorities handle DNSKEY management
+(i.e. the \verb|ch.| zone) and how its authorities handle DNSKEY management
 issues.
@@
 
-By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in fig.~\ref{fig:trust-chain}.
+By signing the keys of the TLD's, the root will put its trust into their keys, and any keys the TLD's have signed, resulting in a trust hierarchy illustrated in figure~\ref{fig:trust-chain}.
 This way, a resolver querying the LACAL's DNS server for \verb|lacal.epfl.ch.| will be able to verify the received query answer against the known root-authority in the following way:
 
+%% TODO this is not entirely true. The DNSKEYs are in the zone
+%% itself. Only their signatures are in the parent zone.
 \begin{enumerate}
@@
 
+%% TODO you cannot talk about TARs here without explaining. Reference
+%% for this?
 DNSSEC's trust hierarchy is similar to the SSL Certificate infrastructure, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public key with additional metadata.}.
@@
   \includegraphics[width=\linewidth]{trust-chain}
-  \caption{Example of a trust-hierarchy. Here, the resolver can establish a trust chain between the Root and the ucla.edu zone.}
+  \caption{Example of a trust-hierarchy. Here, the resolver can establish a trust chain between the Root and the \texttt{ucla.edu.} zone.}
   \label{fig:trust-chain}
@@
 
-\paragraph{Enabling DNSSEC} When enabling DNSSEC for a zone, the
-parent zone has to establish and sign a DS record with the newly
-created public-key.
+\paragraph{Enabling DNSSEC} When enabling DNSSEC for a zone, after it
+has been signed, the parent zone has to establish and sign a DS record
+with the newly created public-key.
 
@@
 irrelevant here, it is important to know that the parent zone has to
-update its DS record regularly.
+update its DS record regularly and hence has to be signed again, too.
 
 \paragraph{Disabling DNSSEC} An operator might choose to discontinue
-DNSSEC for a zone. The DS records (or other means of signing DNSKEYs
-as we will see later) need to be removed.
+DNSSEC for a zone. The DS records in the parent zone (or other means
+of signing DNSKEYs as we will see later) need to be removed.
 
@@
 going into the specific DNSKEY management issues. Please refer to
-section~\ref{sec:case-study} for more details on the situation of the
+section~\ref{sec:case-study} for more details on the situation in the
 \verb|ch.| zone.
@@
 In figure~\ref{fig:dnssec-growth} you can see the development of
-deployed, production DNSSEC zones. The data comes from SecSpider
-\cite{secspider} which crawls as many DNSSEC zones as possible,
-verifies whether they are properly signed and then applies a heuristic to
-determine whether it is actually a production zone. The goal is to
-rule out zones which are created for testing such as
+deployed, production DNSSEC zones in time. The data comes from
+SecSpider \cite{secspider} which crawls as many DNSSEC zones as
+possible, verifies whether they are properly signed and then applies a
+heuristic to determine whether it is actually a production zone. The
+goal is to rule out zones which are created for testing such as
 \url{bogussig.bogussig.test.jelte.nlnet.labs.nl}.
@@
 unable to find an estimate of the number of (production) DNS zones in
-the Internet to estimate the percentage of DNSSEC penetration.
+the Internet in order estimate the percentage of DNSSEC penetration.
 
@@
 Another type of TAR resides at the resolver: A statically configured
-TAR stores keys locally at the resolver and hence needs not to issue an
+TAR stores keys locally at the resolver and hence does not require to issue an
 additional query for keys when a new DNS request arrives. The
 statically configured TAR may automatically poll keys asynchronously,
-or is maybe just a bunch of manually configured trusted keys. As an
+or may be just a bunch of manually configured trusted keys. As an
 example of a statically configured TAR, we have
 Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
+%% TODO what does this "plugin" do exactly?
 to a standard resolver, \cite{Osterweil09} which can poll keys from
@@
 The registries are the entities serving the records for a zone.
-For example, Switch is the registry for the \verb|ch.| zone.
-Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department to be precise \footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
+For example, SWITCH is the registry for the \verb|ch.| zone.
+Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department to be precise\footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
 
+%% TODO description about registrars is wrong. They do not host DNS entries.
 These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a domain name and an IP address.
+Figure~\ref{fig:3Rs} illustrates the relations among the three R's.
 
-In Switzerland, for example, Switch plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from Switch.
+\begin{figure}
+\center
+  \includegraphics[width=\linewidth]{3Rs}
+  \caption{Example of relations between the three R's.}
+  \label{fig:3Rs}
+\end{figure}
+
+In Switzerland, for example, SWITCH plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from SWITCH.
 Registries will generally delegate part or whole of the namespace registration managing to registrars\footnote{For example, one can also buy a .ch domain from \url{http://gandi.net}}
@@
 
+%% TODO again: IP updates (except the Name Servers) don't go through 3R
 The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data was actually updated.
@@
 
-Enabling DNSSEC for a zone not being a \emph{do once and forget} operation, but instead requires additional coordination among the three R's, let us see what the managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.
+%% TODO phrase is a bit clumsy
+Enabling DNSSEC for a zone not being a \emph{do once and forget} operation, but instead requiring additional coordination among the three R's, let us see what managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.
 
@@
 As stated earlier, once the \verb|epfl.ch.| zone is signed it is easy for the school to sign \verb|www.epfl.ch.|.
-However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by Switch or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
+However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by SWITCH or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
 
-While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but Switch fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as Switch's servers do not reflect the update.
+While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but SWITCH fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as SWITCH's servers do not reflect the update.
 
-As suggested by some RFC\cite{RFC4641} key updates should  happen at least every few months.
+As suggested by RFC4641 \cite{RFC4641} key updates should  happen at least every few months.
 Hence, it is crucial that a streamlined and reactive process is available for child-zones to notify their parent-zone when required.
@@
 
+%% TODO add part about computation
+
 \paragraph{Cost} DNSSEC secure delegation is currently provided by

### 2012-12-07 paper/paper.tex
Fix date of root signing in abstract

@@
   contents, especially with respect to the signing of the root zone on
-  July 15, 2012. We will see that DNSSEC penetration in the Internet's 
+  July 15, 2010. We will see that DNSSEC penetration in the Internet's 
   zones is still low, but that the necessary foundations have been lain

### 2012-12-07 paper/paper.tex
Corrections and precisions.

@@
 \section{Introduction}
-DNSSEC is gradually being deployed in the Internet. However, managing
+DNSSEC is gradually being deployed throughout the Internet. However, managing
 the cryptographic keys required for the secure delegation provided by
@@
 The goal of DNSSEC is to allow resolvers -- any person or process querying a DNS server -- to ensure the authenticity of the reply.
-Namely, in standard DNS, nothing guarantees the resolvers that his query's answer has not been modified or really originated from a legitimate server.
+Namely, in standard DNS, nothing guarantees the resolvers that his query's answer has not been modified during the transmission or really originated from a legitimate server.
 
@@
 
-In the DNS world, the \emph{friend} we trust is the so-called root authority -- the root domain -- of whom we know the public key in advance.
+In the DNS world, the \emph{friend} we trust is the so-called root authority -- the root domain -- of whom we know the public key in advance, so we can verify anything it signed.
 
@@
 
-%% TODO this is not entirely true. The DNSKEYs are in the zone
-%% itself. Only their signatures are in the parent zone.
 \begin{enumerate}
-\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s public key, obtained from \verb|epfl.ch.|'s DNS server ;
-\item Verify the query -- containing \verb|lacal.epfl.ch.|'s public key -- returned by \verb|epfl.ch.| with its public key, retrieved from the \verb|ch.| DNS server;
-\item Again, verify \verb|ch.|'s responses by getting its public key from the root domain.
+\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s public key ;
+\item Check \verb|lacal.epfl.ch.|'s public key authenticity by obtaining its signature from \verb|epfl.ch.|'s DNS server ;
+\item Verify the query -- containing \verb|lacal.epfl.ch.|'s public key signature -- returned by \verb|epfl.ch.| with its public key;
+\item Check \verb|epfl.ch.|'s public key against the corresponding signature served by \verb|ch.|;
+\item Again, verify \verb|ch.|'s responses by getting its public key signature from the root domain.
 \item Finally, the root domain's responses can be verified with its already known public key.
@@
 For example, SWITCH is the registry for the \verb|ch.| zone.
-Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department to be precise\footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
+Registrants, on their side, are the entities willing to obtain a certain domain: the registrant of \verb|epfl.ch.| is the school in itself (its IT-department, to be precise\footnote{As a \emph{whois} query on the epfl.ch domain will show you.}), for example.
 
@@
 As stated earlier, once the \verb|epfl.ch.| zone is signed it is easy for the school to sign \verb|www.epfl.ch.|.
-However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.| managed by SWITCH or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well be different among different registrars and TLD's.
+However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.|, managed by SWITCH, or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well vary among different registrars and TLD's.
 
@@
 \begin{description}
-\item[Key Signing] Services that will handle the zone signing and key rollovers are available\footnote{like this registrar: \url{http://registrars.nominet.org.uk/registration-and-domain-management/dns/dnssec-signing-service}}.
+\item[Key Signing] Services that will handle the zone signing and key rollovers are available\footnote{Like this registrar: \url{http://registrars.nominet.org.uk/registration-and-domain-management/dns/dnssec-signing-service}}.
 \item[Hosting] Countless hosting services are available, many of which are proposed by actual registrars.
@@
 \item[-] The transferred private key is not totally \emph{private} anymore, as now both the old and new service provider know it;
-\item[-] Newly generated keys will require an update of the parent zone, and this might happen outside of normally planned rollover periods.
+\item[-] Newly generated keys will require an update of the parent zone
 \end{itemize}
+
+In both cases,  this \emph{transfer} scenario will require a certain level of cooperation between two concurrent companies.
 
@@
 \subsection{Computing power}
-DNSSEC does not only require a higher coordination among the three R's, it also increases the strain on the computing infrastructure of a registrar when the size and quantity of the  zones it manages grow: updates to a zone or key rollovers require cryptographic signing operations that are orders of magnitude more expensive than simply updating the DNS record.
+DNSSEC does not only require a higher coordination among the three R's, it also increases the strain on the computing infrastructure of a registrar when the size and quantity of the  zones it manages grow: updates to a zone or key rollovers require cryptographic signing operations that are orders of magnitude more expensive than simply updating a DNS record.
 Even with improvements like incremental signing, the computational requirements are significant and must be considered when a registrar wishes to enable DNSSEC.
@@
 
-Again, this teardown must happen in a synchronous way -- requiring additional effort with regard to DNS -- as resolvers will perceive the domain to be \emph{broken} as long as all involved parties are in a incoherent state.
+Again, this teardown must happen in a synchronous way -- requiring additional efforts compared to standard DNS -- as resolvers will perceive the domain to be \emph{broken} as long as all involved parties are in an incoherent state.
 

### 2012-12-07 paper/paper.tex
added a ref

@@
 DNSSEC does not only require a higher coordination among the three R's, it also increases the strain on the computing infrastructure of a registrar when the size and quantity of the  zones it manages grow: updates to a zone or key rollovers require cryptographic signing operations that are orders of magnitude more expensive than simply updating a DNS record.
-Even with improvements like incremental signing, the computational requirements are significant and must be considered when a registrar wishes to enable DNSSEC.
+Even with improvements like incremental signing~\cite{signMetrics}, the computational requirements are significant and must be considered when a registrar wishes to enable DNSSEC.
 %TODO find good ref on this.

### 2012-12-07 paper/paper.tex
Minor English fixes

--- 2012-12-07 paper/paper.tex
@@
 
-This is achieved by establishing a trust chain between something known and the domain that is being queried: e.g., I trust this person to tell me the truth and to be who it claims to be, because a friend I trust is trusting this person.
+This is achieved by establishing a trust chain between something known and the domain that is being queried: e.g., I trust this person to tell me the truth and to be who it claims to be, because a friend I trust trusts this person.
 
@@
 \begin{enumerate}
-\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s public key ;
-\item Check \verb|lacal.epfl.ch.|'s public key authenticity by obtaining its signature from \verb|epfl.ch.|'s DNS server ;
+\item Verify the reply's authenticity using \verb|lacal.epfl.ch.|'s
+  public key;
+\item Check the authenticity of \verb|lacal.epfl.ch.|'s public key by
+  obtaining its signature from \verb|epfl.ch.|'s DNS server;
 \item Verify the query -- containing \verb|lacal.epfl.ch.|'s public key signature -- returned by \verb|epfl.ch.| with its public key;
@@
 
-While standard DNS can perfectly live with some of the temporary imprecisions generated by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but SWITCH fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as SWITCH's servers do not reflect the update.
+While standard DNS can perfectly live with some of the temporary imprecisions caused by non-optimal interactions between the three R's, DNSSEC is far less tolerant to such jitter: typically, if the \verb|epfl.ch.| zone updates its keys but SWITCH fails to reflect this change fast enough in its records, query responses from \verb|epfl.ch.|'s DNS servers will appear as illegitimate whereas they perfectly are, and this will last as long as SWITCH's servers do not reflect the update.
 
@@
 
-In both cases,  this \emph{transfer} scenario will require a certain level of cooperation between two concurrent companies.
+In both cases,  this \emph{transfer} scenario will require a certain level of cooperation between two competing companies.
 

### 2012-12-07 paper/paper.tex
new figure for DNSSEC removal

--- 2012-12-07 paper/paper.tex
@@
 
-For DNSSEC's shutdown to be clean and complete, the parent zone and any TAR must be cleaned from whatever DS record they contain for the zone which is being withdrawn from DNSSEC.
+\begin{figure}
+  \includegraphics[width=\linewidth]{trust-chain-tar-no-key}
+  \caption{Illustration of a domain having disabled DNSSEC with a TAR still serving a record for it.}
+  \label{fig:remaining-tar}
+\end{figure}
+
+For DNSSEC's shutdown to be clean and complete, the parent zone and any TAR (figure~\ref{fig:remaining-tar}) must be cleaned from whatever DS record they contain for the zone which is being withdrawn from DNSSEC.
 

### 2012-12-09 paper/paper.tex
first conclusion try...

@@
 
+\section{Conclusion\wbjp\wbts}
+%TODO improve conclusion... :-D
+We have exposed the administrative challenges that arise with DNSSEC: while being far from trivial, these issues are not insurmountable, especially with regard to the gained benefits, and the modifications required from the three R's to their interaction process finally only consists in adaptations to already existing communication channels.
+
+From this point of view, DNSSEC's requirements and goals seem to be reachable.
+
+SWITCH's case study showed that the existing issues can at least be harnessed at a TLD level.
+While a wider adoption is required to see if the infrastructure can indeed scale, the foundations that have been lain seem ready to take on the challenge.
+
 \nocite{*}

### 2012-12-10 paper/paper.tex
Correct conclusion

@@
 \newcommand{\wbts}{\textsuperscript{TS}}
+\newcommand{\wbjpts}{\textsuperscript{JP,TS}}
 
@@
 
-\section{Conclusion\wbjp\wbts}
-%TODO improve conclusion... :-D
-We have exposed the administrative challenges that arise with DNSSEC: while being far from trivial, these issues are not insurmountable, especially with regard to the gained benefits, and the modifications required from the three R's to their interaction process finally only consists in adaptations to already existing communication channels.
+\section{Conclusion\wbjpts}
+We have exposed the administrative challenges that arise with DNSSEC:
+while being far from trivial, these issues are not insurmountable,
+especially with regard to the gained benefits. Further, the
+modifications required from side the three Rs to their interaction
+process finally only consists in adaptations of already existing
+communication channels.
 
-From this point of view, DNSSEC's requirements and goals seem to be reachable.
+From this point of view, DNSSEC's requirements and goals seem to be
+reachable.
 
-SWITCH's case study showed that the existing issues can at least be harnessed at a TLD level.
-While a wider adoption is required to see if the infrastructure can indeed scale, the foundations that have been lain seem ready to take on the challenge.
+SWITCH's case study showed that the existing issues can be harnessed
+at least at a TLD level. While a wider adoption is required to see if
+the infrastructure can indeed scale, the foundations that have been
+lain seem ready to take on the challenge.
 

### 2012-12-10 paper/paper.tex
Fixes for uppercases after colons

@@
 SecSpider \cite{secspider, Osterweil09} is another inline TAR, but
-uses a different approach for key validation: Rather then entering and
+uses a different approach for key validation: rather then entering and
 signing public keys manually, it randomly queries DNSKEY entries from
@@
 
-Another type of TAR resides at the resolver: A statically configured
+Another type of TAR resides at the resolver: a statically configured
 TAR stores keys locally at the resolver and hence does not require to issue an

### 2012-12-10 paper/paper.tex
Chang to en_GB

@@
 \subsubsection{Single administrative domain}
-Basically, once an organization has its main domain signed, it is relatively easy for any sub-entity within said organization to get a sub-domain signed, and to coordinate any operations relative to DNSSEC like key rollover: staying with our EPFL example, it can suffice to contact the IT-department and discuss about the procedure with them\footnote{EPFL does not actually use DNSSEC, but we use the school as an example for the sake of simplicity.}.
-The bottom line is that it is relatively easy to handle the additional work and information updates required by DNSSEC as long as every concerned party lies within the same organization.
+Basically, once an organisation has its main domain signed, it is relatively easy for any sub-entity within said organisation to get a sub-domain signed, and to coordinate any operations relative to DNSSEC like key rollover: staying with our EPFL example, it can suffice to contact the IT-department and discuss about the procedure with them\footnote{EPFL does not actually use DNSSEC, but we use the school as an example for the sake of simplicity.}.
+The bottom line is that it is relatively easy to handle the additional work and information updates required by DNSSEC as long as every concerned party lies within the same organisation.
 
@@
 registered\footnote{\url{https://www.nic.ch/reg/cm/wcm-page/statistics/index.html}}. Upon
-inquiry, SWITCH stated that there are around 370 secure delegations
+enquiry, SWITCH stated that there are around 370 secure delegations
 from the \verb|ch.| zone. This is less than $0.02\%$ penetration

### 2012-12-10 paper/paper.tex
Fix some TODOs in paper

@@
 
-%% TODO you cannot talk about TARs here without explaining. Reference
-%% for this?
 DNSSEC's trust hierarchy is similar to the SSL Certificate infrastructure, where an entity marks its trust in another by \emph{signing} its certificate\footnote{Which is nothing more than a public key with additional metadata.}.
-However, DNSSEC theoretically only has one single top-authority\footnote{The use of TAR's should remain temporary, as in an ideal setup DNSSEC could work without them.}, while the SSL Certificate hierarchy disposes of many.
+However, DNSSEC theoretically only has one single top-authority, while the SSL Certificate hierarchy disposes of many.
 
@@
 example of a statically configured TAR, we have
-Vantages\footnote{\url{http://www.vantage-points.org/}}, an addition
-%% TODO what does this "plugin" do exactly?
-to a standard resolver, \cite{Osterweil09} which can poll keys from
-different locations such as web pages, but also also other
-vantage-enabled resolvers. Of course, the operator may also choose to
-add trusted keys manually.
+Vantages\footnote{\url{http://www.vantage-points.org/}}, a standalone
+daemon, \cite{Osterweil09} which can poll keys from different
+locations such as web pages, but also from other vantage resolvers. Of
+course, the operator may also choose to add trusted keys manually.
 
@@
 
-%% TODO add part about computation
-
 \paragraph{Cost} DNSSEC secure delegation is currently provided by

### 2012-12-10 paper/paper.tex
Todos cleanup

--- 2012-12-10 paper/paper.tex
@@
 
-%% TODO description about registrars is wrong. They do not host DNS entries.
-These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a domain name and an IP address.
+These two entities can eventually be linked by the third one, the registrar: this is where registrants will obtain the actual registration service, namely, the linking between a zone name and their authoritative DNS servers.
 Figure~\ref{fig:3Rs} illustrates the relations among the three R's.
@@
 
-%% TODO again: IP updates (except the Name Servers) don't go through 3R
-The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like IP addresses or billing data was actually updated.
+The entities behind the three R's are the exact same whether DNSSEC is used or not, but DNSSEC modifies the relations among them: e.g., DNSKEY updates will now also require communication between the three R's, even if no practical information like DNS server addresses or billing data was actually updated.
 
@@
 
-%% TODO phrase is a bit clumsy
-Enabling DNSSEC for a zone not being a \emph{do once and forget} operation, but instead requiring additional coordination among the three R's, let us see what managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.
+Enabling DNSSEC for a zone not isn't a \emph{do once and forget} operation, but requires additional coordination among the three R's.
+Let us see what managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.
 
@@
 Even with improvements like incremental signing~\cite{signMetrics}, the computational requirements are significant and must be considered when a registrar wishes to enable DNSSEC.
-%TODO find good ref on this.
 

### 2012-12-10 paper/paper.tex
Some final typo fixes

@@
 daemon, \cite{Osterweil09} which can poll keys from different
-locations such as web pages, but also from other vantage resolvers. Of
+locations such as web pages, but also from other Vantage resolvers. Of
 course, the operator may also choose to add trusted keys manually.
@@
 In Switzerland, for example, SWITCH plays both the roles of the registry and a registrar: other registrars also exist but one can purchase a domain name directly from SWITCH.
-Registries will generally delegate part or whole of the namespace registration managing to registrars\footnote{For example, one can also buy a .ch domain from \url{http://gandi.net}}
+Registries will generally delegate part or whole of the namespace registration management to registrars\footnote{For example, one can also buy a .ch domain from \url{http://gandi.net}}
 
-Registrants are also free to play the registry and registrar roles for their own sub-domains: groups, laboratories and associations within EPFL may apply for an \verb|epfl.ch.| sub-domain, like the LACAL and it's \verb|lacal.epfl.ch.| subdomain, for example.
+Registrants are also free to play the registry and registrar roles for their own sub-domains: groups, laboratories and associations within EPFL may apply for an \verb|epfl.ch.| sub-domain, like the LACAL and its \verb|lacal.epfl.ch.| subdomain, for example.
 
@@
 
-Enabling DNSSEC for a zone not isn't a \emph{do once and forget} operation, but requires additional coordination among the three R's.
-Let us see what managerial consequences arise, depending on if we cross or stay within certain administrative boundaries.
+Enabling DNSSEC for a zone isn't a \emph{do once and forget} operation, but requires additional coordination among the three R's.
+Let us see what managerial consequences arise, depending on whether we cross or stay within certain administrative boundaries.
 
@@
 As stated earlier, once the \verb|epfl.ch.| zone is signed it is easy for the school to sign \verb|www.epfl.ch.|.
-However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLD's (\verb|ch.|, managed by SWITCH, or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have they own terms and conditions, which might very well vary among different registrars and TLD's.
+However, if the domain to be signed was \verb|epfl.ch.| or \verb|epfl.ae|, the school would have to deal with the entity or entities managing the TLDs (\verb|ch.|, managed by SWITCH, or \verb|ae.|\footnote{epfl.ae has been registered at 'Instra Corporation Pty Ltd', a registrar for the ae.}, respectively), which will have their own terms and conditions, that might very well vary among different registrars and TLDs.
 
@@
 \begin{itemize}
-\item[-] The transferred private key is not totally \emph{private} anymore, as now both the old and new service provider know it;
-\item[-] Newly generated keys will require an update of the parent zone
+\item[--] The transferred private key is not totally \emph{private} anymore, as now both the old and new service provider know it;
+\item[--] Newly generated keys will require an update of the parent zone
 \end{itemize}
